---
title: "Module 3 - Describing Data: Central Tendency"
output: html_notebook
---

# Central Tendency
Where the center is at, a statistical measure to determine a single value that defines the center of distribution. Often called averages.
An attempt to find the single value that is most typical or representative of the entire group.
```{r}

# install.packages("RPostgreSQL")
# install.packages("compare")
# install.packages("moments")
require("RPostgreSQL")
require("compare")
require("moments")

drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "cali_water", host = "localhost", port = 5432,user = "ruser", password = "ruser")
sql_string <- paste("SELECT * FROM lab_results WHERE (units = 'mg/L' AND county_name = 'Alameda' AND parameter = 'Dissolved Nitrate')", sep="")
Lab_results <- data.frame(dbGetQuery(con, sql_string))

dbDisconnect(con)
```


# The Mean
The "true" mathematical/arithmetic "average" found by adding all numbers in the data set and then dividing by the number of values in the set. 
```{r}
# average nitrates level per test.
# trim option
mean(Lab_results$result, na.rm = TRUE)
```

# The Median
The median the middle value when a data set is ordered from least to greatest. Appropriate for skewed/open-ended distributions and distributions with undetermined scores.

Bimodel representations indicate that two distinct groups may exist in the same population.
```{r}
# Median nitrate levels
median(Lab_results$result, na.rm = TRUE)
```

# The Mode
The value that occurs most often. Best used for nominal or categorical data.å
```{r}
# One method
# > mode()
# EPA 300.0 28d Hold 
#                764 
mode <- function()
  {
    value <- (sort(-table(Lab_results$method_name))[1])
    return(-value)
  }
 

mode()
# Another Method
# > find_mode(Lab_results$method_name)
# [1] "EPA 300.0 28d Hold"
find_mode <- function(x) 
  {
    u <- unique(x)
    tab <- tabulate(match(x, u))
    u[tab == max(tab)]
  }

find_mode(Lab_results$method_name)

# The third method
# > getMode(Lab_results$method_name)
# [1] "EPA 300.0 28d Hold"
getMode <- function(x) 
  {
    keys <- unique(x)
    keys[which.max(tabulate(match(x, keys)))]
  }

getMode(Lab_results$method_name)

# The fourth?
# > names(table(Lab_results$method_name))[table(Lab_results$method_name)==max(table(Lab_results$method_name))]
# [1] "EPA 300.0 28d Hold"
names(table(Lab_results$method_name))[table(Lab_results$method_name)==max(table(Lab_results$method_name))]

```