---
title: "Module 6 - Describing Data: Probability Sections 1 and 2"
output: html_notebook
---

# Probability II
Probably.
-   Using sample data as a basis to answer questions about a population.
-   Relates samples and populations through populations and proportions.
-   Probability -\> propotion of outcome

$$
\textrm{probability of A} = \frac{\textrm{number of outcomes clased as A}}{\textrm{total number of outcomes}}
$$

Which means I need a dataset with a population.
```{r}
# Load the PostgreSQL driver, create a connection to the postgres database
require("RPostgreSQL")
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "covertype", host = "localhost", port = 5432, user = "ruser", password = "ruser")
sql_string <- paste("SELECT * FROM covertype", sep="")
covertype <- data.frame(dbGetQuery(con, sql_string))
dbDisconnect(con)
rm(con)
```

# Probability and porportions

So let's find the probability of one tree type being 'selected'
```{r}
Area_Subset <- subset(covertype, covertype$wilderness_area4=="TRUE")
uniquetypes <- unique(covertype$cover_type)
treecount <- c(table(sort(covertype$cover_type)))

for(i in uniquetypes)
  {
      print(paste("Probability of selecting CoverType", i, " from Wilderness_Area4 is: ", (sum(treecount[i])/sum(treecount[1:7]))*100, sep=""))
  }
```

Frequency Graphs:
```{r}
Area_Subset <- subset(covertype, covertype$wilderness_area4=="TRUE")
uniquetypes <- unique(covertype$cover_type)
treecount <- c(table(covertype$cover_type))
percentages <- c()
options(warn=-1)

for(i in sort(uniquetypes))
  {
    percentages <- append(percentages, sum(treecount[i])/sum(treecount[1:7])*100)
  }

par(mfrow=c(1,2))
barplot(treecount, main = "Simple Frequency")
barplot(percentages, main = "Probability Frequency?")

par(mfrow=c(2,2))
plot(dnorm(percentages), type = "l")
plot(pnorm(percentages), type = "l")
plot(qnorm(percentages), type = "l")
plot(rnorm(percentages), type = "l")

# Or should that be counts?
par(mfrow=c(1,1))
plot(rnorm(treecount), type = "l")

# If so only rnorm works for treecount and rnorm: "....is used to generate random numbers whose distribution is normal. It takes the sample size as input and generates that many random numbers. We draw a histogram to show the distribution of the generated numbers...."
# Making dnorm the only one that looks like the classic 'normal distribution'.
# But this all feels like incorrect categories...These are all counts and not values. What's the standard deviation for "yes" or "no"? That doesn't seem right. 
options(warn=0)
``` 
CoverType
1 - Spruce/Fir
2 - Lodgepole Pine
3 - Ponderosa Pine
4 - Cottonwood/Willow
5 - Aspen
6 - Douglas-fir
7 - Krummholz

WildernessArea
1 - Rawah Wilderness Area
2 - Neota Wilderness Area
3 - Comanche Peak Wilderness Area
4 - Cache la Poudre Wilderness Area

# Notations and Random Sampling
  Expressed as fractions, decimals, or percentages. 

# The Unit Normal Table
  - Gives the probability that a standard normal random variable Z is less than any given number z. 
  - Used to find the probability that a statistic is observed below, above, or between values on the standard normal distribution, and by extension, any normal distribution.
    
# Probability II
    
# The Binomial Distribution

# Multiplication and Addition Rules


