---
title: "Module 6 - Describing Data: Probability Sections 1 and 2"
output: html_notebook
---

# Probability II

Probably. - Using sample data as a basis to answer questions about a population. - Relates samples and populations through populations and proportions. - Probability -\> propotion of outcome

$$
\textrm{probability of A} = \frac{\textrm{number of outcomes clased as A}}{\textrm{total number of outcomes}}
$$

Which means I need a dataset with a population.

```{r}
# Load the PostgreSQL driver, create a connection to the postgres database
require("RPostgreSQL")
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "covertype", host = "localhost", port = 5432, user = "ruser", password = "ruser")
sql_string <- paste("SELECT * FROM covertype", sep="")
covertype <- data.frame(dbGetQuery(con, sql_string))
dbDisconnect(con)
rm(con)
```

# Probability and porportions

So let's find the probability of one tree type being 'selected'

```{r}
Area_Subset <- subset(covertype, covertype$wilderness_area4=="TRUE")
uniquetypes <- unique(covertype$cover_type)
treecount <- c(table(sort(covertype$cover_type)))

for(i in uniquetypes)
  {
      print(paste("Probability of selecting CoverType", i, " from Wilderness_Area4 is: ", (sum(treecount[i])/sum(treecount[1:7]))*100, sep=""))
  }
```

Frequency Graphs:

```{r}
Area_Subset <- subset(covertype, covertype$wilderness_area4=="TRUE")
uniquetypes <- unique(covertype$cover_type)
treecount <- c(table(covertype$cover_type))
percentages <- c()
options(warn=-1)

for(i in sort(uniquetypes))
  {
    percentages <- append(percentages, sum(treecount[i])/sum(treecount[1:7])*100)
  }

par(mfrow=c(1,2))
barplot(treecount, main = "Simple Frequency")
barplot(percentages, main = "Probability Frequency?")

par(mfrow=c(2,2))
plot(dnorm(percentages), type = "l")
plot(pnorm(percentages), type = "l")
plot(qnorm(percentages), type = "l")
plot(rnorm(percentages), type = "l")

# Or should that be counts?
par(mfrow=c(1,1))
plot(rnorm(treecount), type = "l")

# If so only rnorm works for treecount and rnorm: "....is used to generate random numbers whose distribution is normal. It takes the sample size as input and generates that many random numbers. We draw a histogram to show the distribution of the generated numbers...."
# Making dnorm the only one that looks like the classic 'normal distribution'.
# But this all feels like incorrect categories...These are all counts and not values. What's the standard deviation for "yes" or "no"? That doesn't seem right. 
options(warn=0)
```

CoverType

1.  Spruce/Fir

2.  Lodgepole Pine

3.  Ponderosa Pine

4.  Cottonwood/Willow

5.  Aspen

6.  Douglas-fir

7.  Krummholz

Wilderness area name:

1.  Rawah

2.  Neota

3.  Comanche Peak

4.  Cache la Poudre Wilderness Area

# Notations and Random Sampling

Expressed as fractions, decimals, or percentages.

# The Unit Normal Table

-   Gives the probability that a standard normal random variable Z is less than any given number z.
-   Used to find the probability that a statistic is observed below, above, or between values on the standard normal distribution, and by extension, any normal distribution.

# Probability II

# The Binomial Distribution

-   When a variable is measured in a scale of two categories the resulting data is termed 'binomial'.
-   Off on, yes no.

**dbinom** (probability): returns the value of the probability density function

    dbinom(x,             # X-axis values (x = 0, 1, 2, ..., n)
         size,            # Number of trials (n > = 0)
         prob,            # The probability of success on each trial
         log = FALSE)     # If TRUE, probabilities are given as log

```{r}
#So how does this fit in with the trees?
treecount <- c(table(sort(covertype$cover_type)))
chance <- (sum(treecount[2])/sum(treecount[1:7]))
dbi<-dbinom(x = 1000:2000, size = treecount, prob = chance)

plot(dbi, type = "h", lwd = 1,     
     main = "Ranged 1000 to 2000, how likely are you to have x covertype? \n dbinom()",
     ylab = "P(X = x)", xlab = "X number of trees")

FileDirectory <- paste("/home/daiten/Programming/R/Projects/Biostatistics Tutorial/Media/", sep="")
png(paste(FileDirectory, "TreeProbability.jpg", sep = ""), width = 1000, height = 500)
plot(dbi, type = "h", lwd = 1,     
     main = "Ranged 1000 to 2000, how likely are you to have x covertype? \n dbinom()",
     ylab = "P(X = x)", xlab = "X number of trees")
dev.off()
```

-   Employing dbinom() Finds the probability of getting a certain number of (x) in a certain sample of (size) where (prob) of success on each trial is fixed
-   So, I should be able to graph this: "Probability of selecting X CoverType2s from Wilderness_Area4 with a chance of 48.7599223423957%"
-   Binomial probability mass function (Probability function)
-   When we are interested in knowing the probability of exactly x successes in n trials that has a probability of success, p, on each trial.

**pbinom** (distribution): returns the value of the cumulative density function

    pbinom(q,             # Quantile or vector of quantiles
       size,              # Number of trials (n > = 0)
       prob,              # The probability of success on each trial
       lower.tail = TRUE, # If TRUE, probabilities are P(X <= x), or P(X > x) otherwise
       log.p = FALSE)     # If TRUE, probabilities are given as log

-   Returns the area to the left of a (q) in the binomial distribution.
-   Binomial distribution (Cumulative distribution function)

**qbinom** (quantile): returns the value of the inverse cumulative density function - Binomial quantile function

    qbinom(p,                 # Probability or vector of probabilities
           size,              # Number of trials (n > = 0)
           prob,              # The probability of success on each trial
           lower.tail = TRUE, # If TRUE, probabilities are P(X <= x), or P(X > x) otherwise
           log.p = FALSE)     # If TRUE, probabilities are given as log

-   Given a probability or a set of probabilities, obtains the corresponding binomial quantile.
-   Find the pth quantile of a binomial distribution.

**rbinom** (random) - function allows drawing nn random samples from a binomial distribution in R. - Binomial pseudorandom number generation

    rbinom(n,    # Number of random observations to be generated
           size, # Number of trials (> = 0)
           prob) # The probability of success on each trial

-   generates a vector of binomial distributed random variables given a vector length n, number of trials (size) and probability of success on each trial (prob).
-   draw nn random observations from a binomial distribution in R.

# Multiplication and Addition Rules
